---
title: "A Meta Comparison of R Comparison Functions"
author: "Brodie Gaslam"
output:
    function(...) rmarkdown::html_vignette(..., md_extensions="-markdown_in_html_blocks", css=c(file.path(system.file(package="diffobj"), "css", "diffobj.css"), "styles.css")):
        toc: true

vignette: >
  %\VignetteIndexEntry{diffobj}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}

---

```{r, echo=FALSE}
old.opt <- options(
  diffobj.disp.width=80, diffobj.pager="off", diffobj.format="html"
)
```

Do you care about mean relative differences? A comparison of #rstats comparison
functions.

I wrote `diffobj` to answer a simple question: how are these two R objects
different.  There are many functions available in R and in add-on packages that
address several aspects of the problem, but none that consistently provides me
with the "aha, that's how they're different".

## The Functions

* `all.equal`, from base R
* `testthat::compare`
* `compare::compare`
* `diffobj::diffPrint`

## Vectors

```{r, results="asis"}
all.equal(1:10, c(1:9, 11))
compare::compare(1:10, c(1:9, 11))
testthat::compare(1:10, c(1:9, 11))
diffPrint(1:10, c(1:9, 11))
```

## Data Frames and Matrices

## Lists and Other Objects

